Seagrass mapping and monitoring \\ along the coasts of Crete, Greece

### 2011-02-17 MScDiplom_Lemenkova.tex
30 Comment, 1 Research Question updated

\phantomsection	
\addcontentsline{toc}{section}{Front Page}
\begin{titlepage}
	\centering
	\title{Seagrass mapping and monitoring \\ along the coasts of Crete, Greece}
	\author[1]{by Lemenkova Polina}
	\affil[1]{MSc student at the \textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
	\date{\today}
	\maketitle	
\end{titlepage}
\pagebreak

@@

@@

\subsection{Hypotheses}\index[concepts]{Research!questions!hypothesis}
A statistical testing will be used to compare between the spectral responses of the different seafloor cover types (i.e. sand and \textit{P. oceanica}),
 whether it is spectrally distinct and at least one pair is statistically different at every spectral
band.\vspace{1ex}\\
For the \emph{Research Question 1} the Hypothesis $H_o$ claims: seagrass types are not spectrally distinct from other seafloor types with
varying \textit{in-situ} conditions, which means $H_o: \mu1 =\mu2 =\mu3=...= \mu n$.
The alternative Hypothesis $H_a$ claims the opposite statement: seagrass is spectrally
distinct with varying \textit{in-situ} conditions, $H_o:\mu1\neq\mu2\neq\mu3\neq...\neq\mu n$.\vspace{1ex}\\
For the \emph{Research Question 2} the Hypothesis $H_o$ claims: broadband and hyperspectral remote sensing data can be used for the mapping of \textit{P.oceanica}, because they do not provide enough radiometric information to discriminate sand from seagrass, which means $H_o:\mu1=\mu2=\mu3=...=\mu n$. The alternative Hypothesis $H_a$ claims the opposite statement: broadband and hyperspectral sensors do prove to provide enough radiometric information to discriminate sand from seagrass, 
$H_o: \mu1\neq\mu2\neq\mu3\neq ... \neq\mu n$.\vspace{1ex}\\
The distribution of the spectral responses at every spectral band is assumed to be normal as well as
the equality of the statistical variances.
The hypothesis testing is suggested to be carried out using the \ac{ANOVA} statistical test. The purpose of
\ac{ANOVA} test is to visualize in an effective and quick way the spectral differences between seagrass
species and their spatial distribution. Thus, the key hypotheses of the research will be tested to prove
whether the results of the research are accurate, reasonable and correct.

@@

at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}). 

@@

The method of \textit{least squares}\index[concepts]{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made. We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}, \ac{OLS}: Fig.\ref{tab:15}\label{page-OLS}\\

@@

The \textit{k-means cluster analysis}\index[concepts]{Data!statistical analysis!k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.

@@

areas of quartiles\index[concepts]{Data!statistical analysis!quartiles Q1-Q3} (Q1 - Q3, Fig.\ref{fig:A.30}\label{page-A30}, and shaded Q-Q areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution (spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in
spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the

@@

\paragraph{ANOVA-testing}\index[concepts]{Data!statistical analysis!ANOVA}
We applied \ac{ANOVA} (ANalysis Of Variance) testing for the analysis of the probability of the
spectral reflectance of \textit{P.oceanica} to be more or less spectrally distinct from other seafloor cover
types with changing depth. W are checking the statement Hypothesis $H_o$, is true, which is “the
spectral distinguishability of the seagrass \textit{P.oceanica} from other seafloor types is not changing with
varying \textit{in-situ} conditions, $H_o: \mu1 =\mu2 =\mu3=...= \mu n$, which gives the following outcome.
P(reject $H_o\mid H_o$ is valid)= P(X$>c\mid p$=.05).
where c comes for critical values and p=probability.
The result P=.05 is (Tab.\ref{tab:8}), hence, very small, which makes the statement of Hypothesis $H_o$ highly unlikely
(less than 1 in a 10 chance). The one-way \ac{ANOVA} highlighted a significant difference between data
of spectral reflectances of \textit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
Hypothesis $H_a$=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing with
varying \textit{in-situ} conditions, i.e. increasing depth, $H_o: \mu\neq\mu2\neq\mu3\neq ... \neq\mu n$.
As a result of the statistical testing, we came to the following conclusion.
The Hypothesis $H_a$ is true, which claims that the spectral discernibility of the seagrass \textit{P.oceanica} is
distinctly changing with varying \textit{in-situ} conditions, i.e. increasing depth, $H_o: \mu1\neq\mu2\neq\mu3\neq...\neq\mu n$ .
Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
(Box Plots).
The graphs were plotted on the basis of the following statistical data calculated from the sampling
measurements data: Table 2.

@@

%%% Gretl start
\begin{table}[H]
	\rowcolors{1}{Thistle}{MintCream}
	\begin{center}
		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. \href{http://gretl.sourceforge.net/}{Gretl}­}\label{tab:16}
		Summary Statistics, using the observations 301--325\\

@@

\begin{table}[H]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|}
	\hline\hline
		& \multicolumn{2}{|c|}{Cluster} &  & \multicolumn{2}{|c|}{Cluster}\\
		 & 1 & 2 &  & 1 & 2 \\ \hline\hline
		V2  &	0.265141 &	0.046966 &	V16 &	0.181395 &	0.033191 \\ \hline
		V3 &	0.202643 &	0.036088 &	V17 &	0.202698 &	0.036614 \\ \hline
		V4 &	0.20699 &	0.037789 &	V18 &	0.149364 &	0.028217 \\ \hline
		V5 &	0.298975 &	0.052158 &	V19 &	0.225452 &	0.040193 \\ \hline
		V6 &	0.25915 &	0.045968 &	V20 &	0.135163 &	0.025506 \\ \hline
		V7 &	0.293508 &	0.049314 &	V21 &	0.175176 &	0.031866 \\ \hline
		V8 &	0.193292 &	0.036402 &	V22 &	0.142295 &	0.027148 \\ \hline
		V9 &	0.215848 &	0.038738 &	V23 &	0.19514 &	0.034553 \\ \hline
		V10 &	0.147798 &	0.027489 &	V24 &	0.214946 &	0.038321 \\ \hline
		V11 &	0.188955 &	0.034031 &	V25 &	0.14585 &	0.027665 \\ \hline
		V12 &	0.157947 &	0.029319 &	V26 &	0.134787 &	0.024912 \\ \hline
		V13 &	0.165877 &	0.029961 &	V27 &	0.149217 &	0.028326 \\ \hline
		V14 &	0.137149 &	0.02626 &	V28 &	0.193388 &	0.034999 \\ \hline
		V15 &	0.140563 &	0.026211 &	V29 &	0.214436 &	0.038433 \\ \hline
		 &	 &	 &	V30	 & 0.299183	 & 0.053451 \\ \hline
	\end{tabular}
	\end{center}
	\label{tab:23}
\end{table}
\pagebreak

@@

%%%Gretl begin
\begin{table}[H]
	\rowcolors{1}{Gainsboro}{MintCream}
	\begin{center}
	\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica}. \href{http://gretl.sourceforge.net/}{Gretl}}
		Model 1: WLS, using observations 1--95\\
		Dependent variable: v20\\
		Variable used as weight: v12

@@

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Mean dependent var &  0.136154 & S.D. dependent var &  0.149193 \\
		Sum squared resid &  5.689669 & S.E. of regression &  0.247344 \\
	\end{tabular}
	\label{tab:27}
	\end{center}
\end{table}
%%%Gretl end

@@

\begin{table}[H]\footnotesize
	\caption{Available broadband Landsat satellite images covering the research area of Crete island. Page: $<$\pageref{page-30}$>$}\label{tab:34}
	\centering
	\rowcolors{1}{Ivory}{GhostWhite}
	 \begin{tabular}{| p{1cm} | p{3cm} | p{2cm} | p{5cm} |}
	    \hline
		    \textbf{No} & \textbf{Image cource} & \textbf{Data} & \textbf{Name} \\ \hline \hline
		     1 & Landsat ETM+ & 2005/May/04 & WRS2p181r035L71181035-035-20050504-ETM-GLS2005 \\ \hline
		     2 & Landsat TM & 2006/Nov/07 & WRS2p181r036L5181036-036-20061107-TM-GLS2005 \\ \hline
		     3 & Landsat ETM+ & 2005/Apr/25 & WRS2p182r035L71182035-035-20050425-ETM-GLS2005 \\ \hline
		     4 & Landsat ETM+ & 2000/Jul/09 & WRS2p181r036-7dx-20000709-ETM-GLS2000 \\ \hline
		     5 & Landsat TM & 1987/Jun/10 & LandsatWRS2p183r035p183r035-5dx-19870610-TM-GLS1990 \\ \hline
		     6 & Landsat ETM+/ Earth Sat & 1999/Aug/08 & 071-261Mosaic-LandsatN-35N-35-35ETM-EarthSat-MrSID-19990808-20020624 \\ \hline
		     7 & Landsat ETM+/ Earth Sat & 1999/Aug/08 & 071-260Mosaic-LandsatN-35N-35-30ETM-EarthSat-MrSID-19990808-20020617 \\ \hline
		     8 & Landsat ETM+ & 2000/Jun/30 & WRS2p182r036-7x-20000630-ETM-EarthSat \\ \hline
		     9 & Landsat MSS / Earth Sat & 1975/Jul/26 & LandsatWRS1p196r35-2m-19750726-MSS-EarthSat \\ \hline
		    10 & Landsat TM / Earth Sat & 1987/Jun/10 & 012-807LandsatWRS2p183r35-5t-19870610-TM-EarthSat \\ \hline
		    11 & Landsat ETM+ & 2000/Jun/30 & LandsatWRS2p182r036-7dx-20000630-ETM-GLS2000 \\ \hline
		    12 & Landsat ETM+ & 2005/Apr/09 & LandsatWRS2p182r036L71182036-036-20050409-ETM-GLS2005 \\ \hline
	  \end{tabular}	
\end{table}
\pagebreak
\newpage

>>>>

\phantomsection	
\addcontentsline{toc}{section}{Front Page}
\begin{titlepage}
	\centering
	\title{Seagrass mapping and monitoring \\ along the coasts of Crete, Greece}
	\author[1]{by Lemenkova Polina}
	%\affil[1]{MSc student at the \textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
	\date{\today}
	\maketitle	
\end{titlepage}
\pagebreak

@@

	\acro{WLS}{\href{http://en.wikipedia.org/wiki/Weighted_least_squares}{Weighted Least Squares}}

@@

\subsection{Hypotheses}\index[concepts]{Research!questions!hypothesis}
A statistical testing will be used to compare between the spectral responses of the different seafloor cover types (i.e. sand and \textit{P. oceanica}),
 whether it is spectrally distinct and at least one pair is statistically different at every spectral
band.\vspace{1ex}\\
For the \emph{Research Question 1} the Hypothesis $H_o$ claims: seagrass types are not spectrally distinct from other seafloor types with
varying \textit{in-situ} conditions, which means $H_o: \mu1 =\mu2 =\mu3=...= \mu n$.\\
The alternative Hypothesis $H_a$ claims the opposite statement: seagrass is spectrally
distinct with varying \textit{in-situ} conditions, $H_o:\mu1\neq\mu2\neq\mu3\neq...\neq\mu n$.\vspace{1ex}\\
For the \emph{Research Question 2} the Hypothesis $H_o$ claims: broadband and hyperspectral remote sensing data cannot be used for the mapping of \textit{P.oceanica}, because they do not provide enough radiometric information to discriminate sand from seagrass, which means $H_o:\mu1=\mu2=\mu3=...=\mu n$. \\ The alternative Hypothesis $H_a$ claims the opposite statement: broadband and hyperspectral sensors do prove to provide enough radiometric information to discriminate sand from seagrass, $H_o: \mu1\neq\mu2\neq\mu3\neq ... \neq\mu n$.\vspace{1ex}, and can therefore, be used for the seagrass mapping.\\
The distribution of the spectral responses at every spectral band is assumed to be normal as well as the equality of the statistical variances.
The hypothesis testing is suggested to be carried out using the \ac{ANOVA} statistical test. The purpose of
\ac{ANOVA} test is to visualize in an effective and quick way the spectral differences between seagrass
species and their spatial distribution. Thus, the key hypotheses of the research will be tested to prove
whether the results of the research are accurate, reasonable and correct.

@@

at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}, Tab.\ref{tab-16}).\label{page-sumstat} 

@@

The method of \textit{least squares}\index[concepts]{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made. We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, \ac{WLS}: Tab.\ref{tab:27}\label{page-WLS}, \ac{OLS}: Fig.\ref{tab:15}\label{page-OLS}\\

@@

The \textit{k-means cluster analysis}\index[concepts]{Data!statistical analysis!k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23})\label{page-fincc} in which each observation belongs to the cluster with the nearest mean.

@@

areas of quartiles\index[concepts]{Data!statistical analysis!quartiles Q1-Q3} (Q1 - Q3, Fig.\ref{fig:A.30}\label{page-A30}, and shaded Q-Q areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution (spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the

@@

\paragraph{ANOVA-testing}\index[concepts]{Data!statistical analysis!ANOVA}
We applied \ac{ANOVA} (ANalysis Of Variance) testing for the analysis of the probability of the
spectral reflectance of \textit{P.oceanica} to be more or less spectrally distinct from other seafloor cover
types with changing depth. W are checking the statement Hypothesis $H_o$, is true, which is “the
spectral distinguishability of the seagrass \textit{P.oceanica} from other seafloor types is not changing with
varying \textit{in-situ} conditions, $H_o: \mu1 =\mu2 =\mu3=...= \mu n$, which gives the following outcome.
P(reject $H_o\mid H_o$ is valid)= P(X$>c\mid p$=.05).
where c comes for critical values and p=probability.
The result P=.05 is (Tab.\ref{tab:8}), hence, very small, which makes the statement of Hypothesis $H_o$ highly unlikely
(less than 1 in a 10 chance). The one-way \ac{ANOVA} highlighted a significant difference between data
of spectral reflectances of \textit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
Hypothesis $H_a$=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing, 
and seagrass can be spectrally discriminated from carbonate sand with varying \textit{in-situ} conditions,
 i.e. increasing depth, $H_o: \mu\neq\mu2\neq\mu3\neq ... \neq\mu n$.
As a result of the statistical testing, we came to the following conclusion.
The Hypothesis $H_a$ is true, which claims that the spectral discernibility of the seagrass \textit{P.oceanica} is
distinctly changing and can be discriminated from from other seafloor cover types (carbonate sand) with varying \textit{in-situ} conditions, 
i.e. increasing depth, $H_o: \mu1\neq\mu2\neq\mu3\neq...\neq\mu n$, which positively answers first research question.
Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
(Box Plots).
The graphs were plotted on the basis of the following statistical data calculated from the sampling
measurements data: Table 2.

@@

%%% Gretl start
\begin{table}[H]
	\rowcolors{1}{Thistle}{MintCream}
	\begin{center}
		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. \href{http://gretl.sourceforge.net/}{Gretl}. Page: $<$\pageref{page-sumstat}$>$­}\label{tab:16}
		Summary Statistics, using the observations 301--325\\

@@

\begin{table}[H]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}. Page: $<$\pageref{page-fincc}$>$}\label{tab:23}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|}
	\hline\hline
		& \multicolumn{2}{|c|}{Cluster} &  & \multicolumn{2}{|c|}{Cluster}\\
		 & 1 & 2 &  & 1 & 2 \\ \hline\hline
		V2  &	0.265141 &	0.046966 &	V16 &	0.181395 &	0.033191 \\ \hline
		V3 &	0.202643 &	0.036088 &	V17 &	0.202698 &	0.036614 \\ \hline
		V4 &	0.20699 &	0.037789 &	V18 &	0.149364 &	0.028217 \\ \hline
		V5 &	0.298975 &	0.052158 &	V19 &	0.225452 &	0.040193 \\ \hline
		V6 &	0.25915 &	0.045968 &	V20 &	0.135163 &	0.025506 \\ \hline
		V7 &	0.293508 &	0.049314 &	V21 &	0.175176 &	0.031866 \\ \hline
		V8 &	0.193292 &	0.036402 &	V22 &	0.142295 &	0.027148 \\ \hline
		V9 &	0.215848 &	0.038738 &	V23 &	0.19514 &	0.034553 \\ \hline
		V10 &	0.147798 &	0.027489 &	V24 &	0.214946 &	0.038321 \\ \hline
		V11 &	0.188955 &	0.034031 &	V25 &	0.14585 &	0.027665 \\ \hline
		V12 &	0.157947 &	0.029319 &	V26 &	0.134787 &	0.024912 \\ \hline
		V13 &	0.165877 &	0.029961 &	V27 &	0.149217 &	0.028326 \\ \hline
		V14 &	0.137149 &	0.02626 &	V28 &	0.193388 &	0.034999 \\ \hline
		V15 &	0.140563 &	0.026211 &	V29 &	0.214436 &	0.038433 \\ \hline
		 &	 &	 &	V30	 & 0.299183	 & 0.053451 \\ \hline
	\end{tabular}
	\end{center}	
\end{table}
\pagebreak

@@

%%%Gretl begin
\begin{table}[H]
	\rowcolors{1}{Gainsboro}{MintCream}
	\begin{center}
	\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica}. \href{http://gretl.sourceforge.net/}{Gretl}. Page: $<$\pageref{page-WLS}$>$}\label{tab:27}
		Model 1: WLS, using observations 1--95\\
		Dependent variable: v20\\
		Variable used as weight: v12

@@

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Mean dependent var &  0.136154 & S.D. dependent var &  0.149193 \\
		Sum squared resid &  5.689669 & S.E. of regression &  0.247344 \\
	\end{tabular}
	\end{center}
\end{table}
%%%Gretl end

@@

\begin{table}[H]\footnotesize
	\caption{Available broadband Landsat satellite images covering the research area of Crete Island. Page: $<$\pageref{page-30}$>$}\label{tab:34}
	\centering
	\rowcolors{1}{Ivory}{GhostWhite}
	 \begin{tabular}{| p{1cm} | p{3cm} | p{2cm} | p{5cm} |}
	    \hline
		    \textbf{No} & \textbf{Image cource} & \textbf{Data} & \textbf{Name} \\ \hline \hline
		     1 & Landsat ETM+ & 2005/May/04 & WRS2p181r035L71181035-035-20050504-ETM-GLS2005 \\ \hline
		     2 & Landsat TM & 2006/Nov/07 & WRS2p181r036L5181036-036-20061107-TM-GLS2005 \\ \hline
		     3 & Landsat ETM+ & 2005/Apr/25 & WRS2p182r035L71182035-035-20050425-ETM-GLS2005 \\ \hline
		     4 & Landsat ETM+ & 2000/Jul/09 & WRS2p181r036-7dx-20000709-ETM-GLS2000 \\ \hline
		     5 & Landsat TM & 1987/Jun/10 & LandsatWRS2p183r035p183r035-5dx-19870610-TM-GLS1990 \\ \hline
		     6 & Landsat ETM+/ Earth Sat & 1999/Aug/08 & 071-261Mosaic-LandsatN-35N-35-35ETM-EarthSat-MrSID-19990808-20020624 \\ \hline
		     7 & Landsat ETM+/ Earth Sat & 1999/Aug/08 & 071-260Mosaic-LandsatN-35N-35-30ETM-EarthSat-MrSID-19990808-20020617 \\ \hline
		     8 & Landsat ETM+ & 2000/Jun/30 & WRS2p182r036-7x-20000630-ETM-EarthSat \\ \hline
		     9 & Landsat MSS / Earth Sat & 1975/Jul/26 & LandsatWRS1p196r35-2m-19750726-MSS-EarthSat \\ \hline
		    10 & Landsat TM / Earth Sat & 1987/Jun/10 & 012-807LandsatWRS2p183r35-5t-19870610-TM-EarthSat \\ \hline
		    11 & Landsat ETM+ & 2000/Jun/30 & LandsatWRS2p182r036-7dx-20000630-ETM-GLS2000 \\ \hline
		    12 & Landsat ETM+ & 2005/Apr/09 & LandsatWRS2p182r036L71182036-036-20050409-ETM-GLS2005 \\ \hline
	  \end{tabular}	
\end{table}
\pagebreak
\newpage
